Overrelaxed Sinkhorn--Knopp Algorithm for Regularized Optimal Transport

### 2017-11-22 PosterNIPS2017.tex
Poster (WIP)

\begin{poster}
{
background=plain,
headershade=plain,
headerborder=closed, % Adds a border around the header of content boxes
colspacing=0.8em, % Column spacing
bgColorOne=gris, % Background color for the gradient on the left side of the poster
bgColorTwo=gris, % Background color for the gradient on the right side of the poster
borderColor=CPU, % Border color
headerColorOne=CPU, % Background color for the header in the content boxes (left side)
headerColorTwo=CPU, % Background color for the header in the content boxes (right side)
headerFontColor=white, % Text color for the header text in the content boxes
boxColorOne=white, % Background color of the content boxes
textborder=roundedsmall, % Format of the border around content boxes, can be: none, bars, coils, triangles, rectangle, rounded, roundedsmall, roundedright or faded
eyecatcher=true, % Set to false for ignoring the left logo in the title and move the title left
headerheight=0.10\textheight, % Height of the header
headershape=rounded, % Specify the rounded corner in the content box headers, can be: rectangle, small-rounded, roundedright, roundedleft or rounded
headerfont=\Large\bf\textsc, % Large, bold and sans serif font in the headers of content boxes
%textfont={\setlength{\parindent}{1.5em}}, % Uncomment for paragraph indentation
linewidth=1.8pt, % Width of the border lines around content boxes
columns=2,
}
%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------
%
{\includegraphics[height=5em]{logo_ens_psl.png}} % First university/lab logo on the left
{\bf Overrelaxed Sinkhorn--Knopp algorithm\\ for regularized optimal transport\vspace{0.2em}} 
{{Alexis Thibault, L\'enaïc Chizat, Charles Dossal and  Nicolas Papadakis}} % Author names and institution
{\includegraphics[height=5em]{logo_imb.png}} % Second university/lab logo on the right

@@

\headerbox{Introduction}{name=intro,column=0}{
	Optimal Transport is an efficient and flexible tool to compare two probability distributions. It has been popularized in the computer vision community in the context of discrete histograms [Rubner, 2000]. The introduction of entropic regularization in [Cuturi, 2013] has made possible the use of the fast Sinkhorn--Knopp algorithm (SK) [Sinkhorn, 1964c] scaling with high dimensional data.\\
	\begin{minipage}{\textwidth}
		\centering
		\begin{tcolorbox}[width=0.99\textwidth,/tcb/size=small, /tcb/top=2.0mm]
			\centering
			\begin{tabular}{c c c c}
				\includegraphics[height=3cm]{1d_interp_mu} &
				\includegraphics[height=3cm]{1d_interp_0_2} &
				\includegraphics[height=3cm]{1d_interp_2} &
				\includegraphics[height=3cm]{1d_interp_18} \\
				&
				$\epsilon = 5$ &
				$\epsilon = 0.5$ &
				$\epsilon = 0.02$
			\end{tabular}
			\captionof{figure}{Regularized optimal transport maps for decreasing regularization parameter $\epsilon$.} \label{fig:1d_ot}
		\end{tcolorbox}
	\end{minipage}
	For small regularization $\epsilon$, the SK algorithm becomes extremely slow. To counter this, we propose a new algorithm that accelerates SK by \emph{overrelaxation}.
}

@@

\headerbox{Overrelaxation}{name=overrelaxation,column=1}{
Although the SK algorithm is much more efficient than linear solvers in a high-dimensional setting, it becomes extremely slow when the regularization $\epsilon$ is small.
To counteract this limitation, we use the principle of \emph{overrelaxation}.
\begin{minipage}{0.5\textwidth}
	\begin{tcolorbox}[size=small, top=2mm]
		\centering
		\begin{tabular}{c c}
		\input{schema_b} &
		\input{schema_c} \\
		SK &
		Overrelaxed SK
		\end{tabular}
	\captionof{figure}{For small regularization $\epsilon$, the SK algorithm converges slowly, whereas the overrelaxed SK algorithm converges with a better rate}
	\end{tcolorbox}
\end{minipage}
\begin{minipage}{0.49\textwidth}
	We define the overrelaxed projections of parameter $\theta \in \IR$:
	\begin{align*}\label{or_scaling}
	P_{\Ccal_1}^\theta(\gamma) &= \diag(a)^\theta \gamma\\
	P_{\Ccal_2}^\theta(\gamma) &= \gamma \diag(b)^\theta \nonumber
	\end{align*}
	(element-wise exponentiation)
\end{minipage}
}

@@

\headerbox{Lyapunov Function}{name=lyapunov,below=Sinkhorn,column=0}{

@@

\headerbox{Experiments}{name=expe,below=overrelaxation,column=1}{

@@

\headerbox{References}{name=references,below=lyapunov,span =2}{
\noindent [1] M. Cuturi. Sinkhorn distances: Lightspeed computation of optimal transport. {\it Advances in Neural Information Processing Systems 26}, 2013.\\
...
}

>>>>

\begin{poster}
{
background=plain,
headershade=plain,
headerborder=closed, % Adds a border around the header of content boxes
colspacing=0.8em, % Column spacing
bgColorOne=gris, % Background color for the gradient on the left side of the poster
bgColorTwo=gris, % Background color for the gradient on the right side of the poster
borderColor=CPU, % Border color
headerColorOne=CPU, % Background color for the header in the content boxes (left side)
headerColorTwo=CPU, % Background color for the header in the content boxes (right side)
headerFontColor=white, % Text color for the header text in the content boxes
boxColorOne=white, % Background color of the content boxes
textborder=roundedsmall, % Format of the border around content boxes, can be: none, bars, coils, triangles, rectangle, rounded, roundedsmall, roundedright or faded
eyecatcher=true, % Set to false for ignoring the left logo in the title and move the title left
headerheight=0.10\textheight, % Height of the header
headershape=rounded, % Specify the rounded corner in the content box headers, can be: rectangle, small-rounded, roundedright, roundedleft or rounded
headerfont=\Large\bf\textsc, % Large, bold and sans serif font in the headers of content boxes
%textfont={\setlength{\parindent}{1.5em}}, % Uncomment for paragraph indentation
linewidth=1.8pt, % Width of the border lines around content boxes
columns=2,
}
%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------
%
{\includegraphics[height=6em]{logo_ens_psl.png}} % First university/lab logo on the left
{\bf Overrelaxed Sinkhorn--Knopp algorithm\\ for regularized optimal transport\vspace{0.2em}} 
{{Alexis Thibault, L\'enaïc Chizat, Charles Dossal and  Nicolas Papadakis}} % Author names and institution
{\includegraphics[height=6em]{logo_imb.png}} % Second university/lab logo on the right

@@

\headerbox{Introduction}{name=intro,column=0}{
	Optimal Transport is an efficient and flexible tool to compare two probability distributions. It has been popularized in the computer vision community in the context of discrete histograms [Rubner, 2000]. The introduction of entropic regularization in [Cuturi, 2013] has made possible the use of the fast Sinkhorn--Knopp algorithm (SK) [Sinkhorn, 1964c] scaling with high dimensional data.\\
	\begin{minipage}{\textwidth}
		\centering
		\begin{tcolorbox}[width=0.99\textwidth,/tcb/size=small, /tcb/top=2.0mm]
			\centering
			\begin{tabular}{c c c c}
				\includegraphics[height=3cm]{1d_interp_mu} &
				\includegraphics[height=3cm]{1d_interp_0_2} &
				\includegraphics[height=3cm]{1d_interp_2} &
				\includegraphics[height=3cm]{1d_interp_18} \\
				&
				$\epsilon = 5$ &
				$\epsilon = 0.5$ &
				$\epsilon = 0.02$
			\end{tabular}
			\captionof{figure}{Regularized optimal transport maps for decreasing regularization parameter $\epsilon$.} \label{fig:1d_ot}
		\end{tcolorbox}
	\end{minipage}
	For small regularization $\epsilon$, the SK algorithm becomes extremely slow. To counter this, we propose a new algorithm that accelerates SK by \emph{overrelaxation}.
	This idea has been shown to successfully accelerate the algorithm. However, there is no proof of convergence in the literature, and no automated choice of parameters.
	In our paper, we suggest a way of ensuring global convergence. This is done simply by choosing the overrelaxation parameters at each step of the algorithm so that a given function decreases. We will motivate our choice of function with numerical results.
}

@@

\headerbox{Overrelaxation}{name=overrelaxation,column=1}{
\begin{minipage}{0.49\textwidth}
	\begin{tcolorbox}[size=small, top=2mm]
		\centering
		\begin{tabular}{c c}
		\input{schema_b} &
		\input{schema_c} \\
		SK &
		Overrelaxed SK
		\end{tabular}
	\captionof{figure}{For small regularization $\epsilon$, the SK algorithm converges slowly, whereas the overrelaxed SK algorithm converges with a better rate}
	\end{tcolorbox}
\end{minipage}
\begin{minipage}{0.49\textwidth}
	\vspace{0.6mm}
	Although the SK algorithm is much more efficient than linear solvers in a high-dimensional setting, it becomes extremely slow when the regularization $\epsilon$ is small.
	To counteract this limitation, we use the principle of \emph{overrelaxation}.

	We define the overrelaxed projections of parameter $\theta \in \IR$:
	\begin{align*}\label{or_scaling}
	P_{\Ccal_1}^\theta(\gamma) &= \diag(a)^\theta \gamma\\
	P_{\Ccal_2}^\theta(\gamma) &= \gamma \diag(b)^\theta \nonumber .
	\end{align*}
	In these equations, the exponentiation is applied element-wise.

	In the euclidean setting, which also corresponds to the linearization of the operators around the solution, iterative overrelaxed projections can be shown to converge for any $\theta \in (0,2)$.
\end{minipage}
But in general, there is no proof of global convergence for iterative overrelaxed Bregman projections. In order to ensure convergence, we use a Lyapunov function.
}

@@

\headerbox{Lyapunov Function}{name=lyapunov,column=1,below=overrelaxation,bottomaligned=Sinkhorn}{

@@

\headerbox{Experiments}{name=expe,below=Sinkhorn,column=0,span=2}{

@@

\headerbox{References}{name=references,below=expe,span =2}{
\noindent [1] M. Cuturi. Sinkhorn distances: Lightspeed computation of optimal transport. {\it Advances in Neural Information Processing Systems 26}, 2013.\\
...
}
